---
title: "Sampling Rosario"
author: "Einav Grinberg"
date: "10/18/2021"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(sf)
library(spatstat)
library(raster)
library(units)
library(plotrix)  
library(sfsmisc)
library(mapview)
library(rgdal)
library(knitr)
input_path <- setwd(".../data")
files <- list.files(input_path, pattern = "[.]shp$")
```

# Simple Random Spatial Sampling 

Two sets of random spatial samples have been drawn from the city of Rosario.

1. A sample of 100 building rooftops
2. A sample of 300 building rooftops

The samples were created using the **Vector Research Tools** -> **Random points inside a polygon** in QGIS^[https://www.qgis.org/en/site/]. Then the building rooftops were digitized in QGIS using Google Satellite Hybrid^[https://qms.nextgis.com/geoservices/1135/], a Tile Map Service (TMS) layer.

For every sample the rooftop area $m^2$, the mean global horizontal irradiation $(\frac{kWh}{m^2})$, the usable solar radiation $(kWh)$ and renewable electricity production $(kWh)$ were calculated. 

*Buildings rooftops area that are equal and under 30 $m^2$ are defined as 0. Some of the sample points were computed on non built-up areas/roads/parks etc., therefore they were given a 0 to include a density factor in the calculation.*

```{r random sampling, warning=FALSE, echo=FALSE}
rosario_poly <- read_sf("rosario_b1.shp")
rosario_poly <- st_transform(rosario_poly, "+proj=merc +lon_0=0 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")

sample1 <- read_sf("sample100.shp")
sample1 <- st_transform(sample1, "+proj=merc +lon_0=0 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")
#change to 0 above and equal 30 sqm
sample1$elec_prod[sample1$area <= 30] <- 0

sample1$elec_prod_mwh <- sample1$elec_prod/1000
sample1$elec_prod_gwh <- sample1$elec_prod/1000000

head(sample1)
summary(sample1)

sample2 <- read_sf("sample300.shp")
sample2 <- st_transform(sample2, "+proj=merc +lon_0=0 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")
#change to 0 above and equal 30 sqm
sample2$elec_prod[sample2$area <= 30] <- 0

sample2$elec_prod_mwh <- sample2$elec_prod/1000
sample2$elec_prod_gwh <- sample2$elec_prod/1000000

head(sample2)
summary(sample2)

mapview(rosario_poly,  map.types =  "CartoDB.VoyagerLabelsUnder", alpha.regions = 0, legend=F, color="red", lwd=2) + mapview(sample1, col.regions="blue") + mapview(sample2, col.regions="green")
```

## Sample 1 computations

Calculation of the sample mean is represented by $\overline{y}$.
Calculation of the sample variance is represented by $s$.

```{r sample 1 c, warning=FALSE, echo=FALSE}
boxplot(sample1$elec_prod_mwh, horizontal=TRUE, main="Sample 1", xlab = "Renewable Electricity Production Potential (mWh)")

n1 <- as.numeric(nrow(sample1))
y1 <- mean(sample1$elec_prod_mwh)
paste("The mean of sample 1: ", round(y1,2), "(mWh)")
N_all <- 357126 #all buildings in the city
s1 <- var(sample1$elec_prod_mwh)
paste("The variance of sample 1: ", round(s1,2), "(mWh)")
```
The following equation calculates the unbiased variance of the estimator $\overline{y}$: 

\begin{equation} 
  \hat{var}(\overline{y})= (\frac{N-n}{N})*(\frac{s^2}{n}) (\#eq:unvar)
\end{equation} 

* $N$ is the total population size -> the number of all the buildings in Rosario (357,126). 
* $n$ is the sample size -> 100. 
* $s^2$ is the sample variance. 

The following equation calculates the estimated standard error of the estimator $\overline{y}$: 

\begin{equation} 
  SEM = \sqrt{\hat{var}(\overline{y})} (\#eq:se)
\end{equation} 

```{r variance, warning=FALSE, echo=FALSE}
var_un1 <- ((N_all-n1)/N_all)*(s1/n1)
paste("The variance of the sample mean: ", round(var_un1,2) , "(mWh)")

ese1 = sqrt(var_un1) #estimated standard error of variance
paste("The estimated standard error of the sample mean: ", round(ese1,2), "(mWh)")
```

The following equation calculates an unbiased estimator of the population total $\hat{t}$: 
\begin{equation} 
  \hat{t} = N*{\overline{y}} (\#eq:t)
\end{equation} 

The following equation calculates the unbiased variance of the estimator $\hat{t}$: 

\begin{equation} 
  \hat{var}(\hat{t})= N^2*\hat{var}(\overline{y}) (\#eq:tv)
\end{equation} 

The following equation calculates the estimated standard error of the estimator $\hat{t}$: 
\begin{equation} 
  SEM = \sqrt{\hat{var}(\hat{t})} (\#eq:set)
\end{equation} 

```{r total, warning=FALSE, echo=FALSE}
t_un1 = N_all*y1 
paste("The estimation of the renewable electricity production potential by all the buildings in the city:", round(t_un1,2), "(mWh)")

t_var_un1 = (N_all^2) * var_un1 #unbiased variance with the estimate of the total
paste("The variance of the estimated total:", round(t_var_un1,2), "(mWh)")

ese_t1 = sqrt(t_var_un1) #estimated standard error of total
paste("The estimated standard error of the total:", round(ese_t1,2), "(mWh)")
```

```{r CI 1, warning=FALSE ,echo=FALSE}
#confidence interval with 95% 
CI1 <- t_un1 + qt(c(0.05,0.95), df=99) * ese_t1
paste("The 95% confidence interval estimation for sample 1 is: ", "(",round(CI1[1],2)," (mWh), ",round(CI1[2],2)," (mwh))", sep = "")
```

## Sample 2 computations

Calculation of the sample mean is represented by $\overline{y}$.
Calculation of the sample variance is represented by $s$.

```{r sample 2 c, warning=FALSE, echo=FALSE}
boxplot(sample2$elec_prod_mwh, horizontal=TRUE, main="Sample 2", xlab = "Renewable Electricity Production Potential (mWh)")

n2 <- as.numeric(nrow(sample2))
y2 <- mean(sample2$elec_prod_mwh)
paste("The mean of sample 2: ", round(y2,2), "(mWh)")
N_all <- 357126 #all buildings in the city
s2 <- var(sample2$elec_prod_mwh)
paste("The variance of sample 2: ", round(s2,2), "(mWh)")
```

Equation \@ref(eq:unvar) calculates the unbiased variance of the estimator $\overline{y}$. 

Equation \@ref(eq:se) calculates the estimated standard error of the estimator $\overline{y}$.

```{r variance 2, warning=FALSE, echo=FALSE}
var_un2 <- ((N_all-n2)/N_all)*(s2/n2)
paste("The variance of the sample mean: ", round(var_un2,2), "(mWh)")

ese2 = sqrt(var_un2) #estimated standard error of variance
paste("The estimated standard error of the sample mean: ", round(ese2,2), "(mWh)")
```

Equation \@ref(eq:t) calculates an unbiased estimator of the population total $\hat{t}$.

Equation \@ref(eq:tv) calculates the unbiased variance of the estimator $\hat{t}$.

Equation \@ref(eq:set) calculates the estimated standard error of the estimator $\hat{t}$.

```{r total 2, warning=FALSE, echo=FALSE}
t_un2 = N_all*y2 
paste("The estimation of the renewable electricity production potential by all the buildings in the city:", round(t_un2,2), "(mWh)")

t_var_un2 = (N_all^2) * var_un2 #unbiased variance with the estimate of the total
paste("The variance of the estimated total:", round(t_var_un2,2), "(mWh)")

ese_t2 = sqrt(t_var_un2) #estimated standard error of total
paste("The estimated standard error of the total:", round(ese_t2,2), "(mWh)")
```

```{r CI 2, warning=FALSE ,echo=FALSE}
#confidence interval with 95% 
CI2 <- t_un2 + qt(c(0.05,0.95), df=299) * ese_t2
paste("The 95% confidence interval estimation for sample 2 is: ", "(",round(CI2[1],2)," (mWh), ",round(CI2[2],2)," (mwh))", sep = "")
```

## Sample 1 and 2 combined computations 

The following calculations are for both sample 1 and sample 2 together, resulting in a total sample of 400 building rooftops. We will call this sample, sample 3. 

Calculation of the sample mean is represented by $\overline{y}$.
Calculation of the sample variance is represented by $s$.
```{r both, warning=FALSE, echo=FALSE}
#both samples
sample3 <- rbind(sample1, sample2)

n3 <- as.numeric(nrow(sample3))
y3 <- mean(sample3$elec_prod_mwh)
paste("The mean of sample 3: ", round(y3,2), "(mWh)")
N_all <- 357126 #all buildings in the city
s3 <- var(sample3$elec_prod_mwh)
paste("The variance of sample 3: ", round(s3,2), "(mWh)")
```

Equation \@ref(eq:unvar) calculates the unbiased variance of the estimator $\overline{y}$. 

Equation \@ref(eq:se) calculates the estimated standard error of the estimator $\overline{y}$.

```{r variance 3, warning=FALSE, echo=FALSE}
var_un3 <- ((N_all-n3)/N_all)*(s3/n3)
paste("The variance of the sample mean: ", round(var_un3,2), "(mWh)")

ese3 = sqrt(var_un3) #estimated standard error of variance
paste("The estimated standard error of the sample mean: ", round(ese3,2), "(mWh)")
```

Equation \@ref(eq:t) calculates an unbiased estimator of the population total $\hat{t}$.

Equation \@ref(eq:tv) calculates the unbiased variance of the estimator $\hat{t}$.

Equation \@ref(eq:set) calculates the estimated standard error of the estimator $\hat{t}$.

```{r total 3, warning=FALSE, echo=FALSE}
t_un3 = N_all*y3
paste("The estimation of the renewable electricity production potential by all the buildings in the city:", round(t_un3,2), "(mWh)")

t_var_un3 = (N_all^2) * var_un3 #unbiased variance with the estimate of the total
paste("The variance of the estimated total:", round(t_var_un3,2), "(mWh)")

ese_t3 = sqrt(t_var_un3) #estimated standard error of total
paste("The estimated standard error of the total:", round(ese_t3,2), "(mWh)")
```

```{r CI 3, warning=FALSE ,echo=FALSE}
#confidence interval with 95% 
CI3 <- t_un3 + qt(c(0.05,0.95), df=399) * ese_t3
paste("The 95% confidence interval estimation for sample 3 is: ", "(",round(CI3[1],2)," (mWh), ",round(CI3[2],2)," (mwh))", sep = "")
```

# Stratified Random Spatial Sampling 

Rosario is comprised by 6 districts:

1. The northern district (Distrito Norte)
2. The northwest district (Distrito Noroeste)
3. The western district (Distrito Oeste)
4. The southwest district (Distrito Sudoeste)
5. The southern district (Distrito Sur)
6. The central district (Distrito Centro)

The division of each municipal district in Rosario was framed in a general structuring that contemplated the construction and connection of routes from north to south that allowed to create a new metropolitan axis. The formation of the districts had an important impact on the  urban structure, by connecting the  agriculture and peripheral districts (north, northwest and west districts) with the city. 
This enabled opening large avenues, constructing important road junctions, creating public, community and service spaces, and in some districts building new residential homes.

For each district 20 random points were computed and then 20 building rooftops were digitized using the same methods as in chapter \@ref(simple-random-spatial-sampling).

```{r stratified sampling, warning=FALSE, echo=FALSE}
rosario_blocks <- read_sf("rosario_blocks.shp")
rosario_blocks <- st_transform(rosario_blocks, "+proj=merc +lon_0=0 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")
# files.strat <- grep("sample_g", files, value=TRUE)
# list.strat <- lapply(files.strat, read_sf)

g1 <- read_sf("sample_g1.shp")
g1 <- st_transform(g1, "+proj=merc +lon_0=0 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")
g1$group <- 1
g1$district <- "Distrito Norte"
g2 <- read_sf("sample_g2.shp")
g2 <- st_transform(g2, "+proj=merc +lon_0=0 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")
g2$group <- 2
g2$district <- "Distrito Noroeste"
g3 <- read_sf("sample_g3.shp")
g3 <- st_transform(g3, "+proj=merc +lon_0=0 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")
g3$group <- 3
colnames(g3)[5] <- "X_mean"
g3$district <- "Distrito Centro"
g4 <- read_sf("sample_g4.shp")
g4 <- st_transform(g4, "+proj=merc +lon_0=0 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")
g4$group <- 4
g4$district <- "Distrito Sur"
g5 <- read_sf("sample_g5.shp")
g5 <- st_transform(g5, "+proj=merc +lon_0=0 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")
g5$group <- 5
g5$district <- "Distrito Sudoeste"
g6 <- read_sf("sample_g6.shp")
g6 <- st_transform(g6, "+proj=merc +lon_0=0 +k=1 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")
g6$group <- 6
g6$district <- "Distrito Oesto"

#combine stratas and change to 0 above and equal 30 sqm
strat <- rbind(g1,g2,g3,g4,g5,g6)
strat$elec_prod[strat$area <= 30] <- 0
strat$elec_prod_mwh <- strat$elec_prod/1000
strat$elec_prod_gwh <- strat$elec_prod/1000000

head(strat)
summary(strat)

mapview(rosario_blocks,  map.types =  "CartoDB.VoyagerLabelsUnder", alpha.regions = 0.15, zcol="Name", lwd=2) + mapview(strat, col.regions="red")
```

After stratification, the stratas are combined to one sample and the computations are the same as for random sampling.

```{r strat c, warning=FALSE,echo=FALSE}
boxplot(strat$elec_prod_mwh ~ strat$district, main="Samples from Rosario Districts", ylab = "Renewable Electricity Production Potential (mWh)", xlab = "Strata by District",cex.axis = 0.7)

nstrat <- as.numeric(nrow(strat))
# calculate means for each group
aggregate(data=strat, elec_prod_mwh~district, FUN=mean)

ystrat <- mean(strat$elec_prod_mwh)
paste("The mean of the stratfied sample is: ", round(ystrat,2), "(mWh)")
N_all <- 357126 #all buildings in the city
sstrat <- var(strat$elec_prod_mwh)
paste("The variance of the stratfied sample is: ", round(sstrat,2), "(mWh)")
```

Equation \@ref(eq:unvar) calculates the unbiased variance of the estimator $\overline{y}$. 

Equation \@ref(eq:se) calculates the estimated standard error of the estimator $\overline{y}$.

```{r variance strat, warning=FALSE, echo=FALSE}
var_unstrat <- ((N_all-nstrat)/N_all)*(sstrat/nstrat)
paste("The variance of the sample mean: ", round(var_unstrat,2), "(mWh)")

esestrat = sqrt(var_unstrat) #estimated standard error of variance
paste("The estimated standard error of the sample mean: ", round(esestrat,2), "(mWh)")
```

Equation \@ref(eq:t) calculates an unbiased estimator of the population total $\hat{t}$.

Equation \@ref(eq:tv) calculates the unbiased variance of the estimator $\hat{t}$.

Equation \@ref(eq:set) calculates the estimated standard error of the estimator $\hat{t}$.

```{r total strat, warning=FALSE, echo=FALSE}
t_unstrat = N_all*ystrat
paste("The estimation of the renewable electricity production potential by all the buildings in the city:", round(t_unstrat,2), "(mWh)")

t_var_strat = (N_all^2) * var_unstrat #unbiased variance with the estimate of the total
paste("The variance of the estimated total:", round(t_var_strat,2), "(mWh)")

ese_tstrat = sqrt(t_var_strat) #estimated standard error of total
paste("The estimated standard error of the total:", round(ese_tstrat,2), "(mWh)")
```

```{r CI strat, warning=FALSE ,echo=FALSE}
#confidence interval with 95% 
CIstrat <- t_unstrat + qt(c(0.05,0.95), df=119) * ese_tstrat
paste("The 95% confidence interval estimation for the stratified sample is: ", "(",round(CIstrat[1],2)," (mWh), ",round(CIstrat[2],2)," (mwh))", sep = "")
```
# 95% Confidence Intervals

```{r CI plot, echo=FALSE, warning=FALSE}
table <- data.frame(sample=c("Sample 1 - 100", "Sample 2 - 300", "Both Samples - 400", "Stratified Sample - 120"),
                    mean=c(y1, y2, y3, ystrat), 
                    variance_of_mean=c(var_un1, var_un2, var_un3, var_unstrat),
                    standard_error_mean=c(ese1,ese2,ese3,esestrat), total_estimation=c(t_un1, t_un2, t_un3, t_unstrat), variance_of_total=c(t_var_un1, t_var_un2, t_var_un3, t_var_strat), standard_error_total=c(ese_t1, ese_t2, ese_t3, ese_tstrat), CI_low=c(CI1[1],CI2[1],CI3[1],CIstrat[1]), CI_up=c(CI1[2],CI2[2],CI3[2],CIstrat[2]))

plotCI(y = table$total_estimation,               # plotrix plot with confidence intervals
       x = 1:4, li = table$CI_low,
       ui = table$CI_up ,scol="red", xaxt="n", yaxt="n",
       xlab="", ylab="Renewable Electricity (mWh)", main="Confidence Intervals", xlim=c(1,4))
#text(x = totals$sample,y = totals$value, totals$value, pos = 3,srt=45)
axis(side=1,at=1:4,  ## add custom x-axis
     label=c("sample 1 - 100", "sample 2 - 300", "both samples - 400","stratified sample - 120"), cex.axis = 0.7)
eaxis(2, at.small = F, cex.axis = 0.7)
```

# Summarizing Table

```{r table, warning=FALSE,echo=FALSE}
kable(table, digits = 2, format.args = list(big.mark = ",", scientific = FALSE), col.names = c('Sample', 'Mean', 'Var of Mean', 'SEM', 'Total Estimation', 'Var of Total', 'SET', 'Lower CI', 'Upper CI'), algin="lccccccccc", caption = "Summary of Computation Results (mWh)")
```
